plackett_luce_monster

### 2022-07-06 inst/plackett_luce_monster.Rmd
more thoughts

* Runner $i$ might wish to support $j$.  If runner $i$ finishes, once
  his rank is established he might be able to help $j$ in some way.
  He might cheer his still-running friend.  In general, we can
  consider $n$ ordered pairs $(i_r,j_r)$, $1\leq r\leq n$, with runner
  $i_1$ helping runner $j_1$, runner $i_2$ helping $j_2$, and so on.
  One simple case would be equivalence classes of
  $[n]=\{1,2,\ldots,n\}$ with each equivalence class comprising
  mutually supporting runners.  In this model, with a pair of mutually
  supporting runners only the higher-placed competitor can affect the
  other.  The causation is not symmetrical.
* Runner $i$ might hate runner $j$.  If $i$ finishes, he might expend
  his spare energy hinderig $j$ (by throwing stones, for example).
  Note that equivalence classes of mutual hatred might exist.  Again
  note that the link is not symmetrical.
* The help or hindrance discussed above might be contingent on
  placing: perhaps only the winnner (that is, first-placing
  competitor) would be able to influence other competitors.  Or maybe
  competitors with any placing could influence other competitors.
* Runners $i$ and $j$ might elect to cross the finishing line holding
  hands, with the apparent tie broken randomly.  This would be a
  symmetrical relationship.  In general, we could consider
  hand-holding equivalence classes.
* We might have a resource allocation problem, with runners having a
  certain [possibly unknown] amount of "strength" but also the ability
  to expend this strength on objectives other than ranking in the
  race.  A runner might be able to expend just enough energy to remain
  slightly ahead of a rival, reserving energy for other purposes.  If
  runner $i$ wishes to beat runner $j$, and has sufficient resources
  to do so, then runners $i$ and $j$ will finish consecutively with
  $i$ ahead of $j$.

@@

```{tikz, fig.cap = "Partial probability tree for five competitors $a$-$e$ with mutually supporting subsets $(abc)$ and $(de)$, reified approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$(abc)(de)$};

@@

\node at (6, 5)  {$\frac{b+M}{b+c+d+e+2M}$};
\node at (6, 4)  {$\frac{c+M}{b+c+d+e+2M}$};
\node at (6, 3)  {$\frac{d}{b+c+d+e+2M}$};
\node at (6, 2)  {$\frac{e}{b+c+d+e+2M}$};

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};

\node at (9.3, 6) {$\frac{c+M}{c+d+e+M}$};
\node at (9.3, 5) {$\frac{d}{c+d+e+M}$};
\node at (9.3, 4) {$\frac{e}{c+d+e+M}$};

@@

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{c}{c+e+M}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{e+M}{c+e+M}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{c}{c+e+M}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{e+M}{c+e+M}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+e+M}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+e+M}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+e+M}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{e+M}{a+b+c+e+M}$};

@@

\node at (11.8, 0) {$\frac{b+M}{b+c+2M}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{c+M}{b+c+2M}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{a+M}{a+c+2M}$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{c+M}{a+c+2M}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{a+M}{a+b+2M}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{b+M}{a+b+2M}$};

\node at (9.3, 2) {$\frac{b+M}{b+c+e+2M}$};
\node at (9.3, 1) {$\frac{c+M}{b+c+e+2M}$};
\node at (9.3, 0) {$\frac{e}{b+c+e+2M}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{b+M}{b+e+M}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{e}{b+e+M}$};

```{tikz, fig.cap = "Partial probability tree for five competitors $a$-$e$ with mutually supporting subsets $(abc)$ and $(de)$, hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$(abc)(de)$};

\newcommand{\lb}{\lambda b }
\newcommand{\lc}{\lambda c }
\newcommand{\lee}{\lambda e }
\node at (6, 5)  {$\frac{\lb}{\lb+\lc+d+e}$};
\node at (6, 4)  {$\frac{\lc}{\lb+\lc+d+e}$};
\node at (6, 3)  {$\frac{d  }{\lb+\lc+d+e}$};
\node at (6, 2)  {$\frac{e  }{\lb+\lc+d+e}$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{c}{c+\lee}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{\lee}{c+\lee}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{c}{c+\lee}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{\lee}{c+\lee}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+\lee}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+\lee}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+\lee}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{\lee}{a+b+c+\lee}$};

\node at (11.8, 0) {$\frac{b}{b+c}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{c}{b+c}$};

@@

@@

>>>>

* Runner $i$ might wish to support $j$.  If runner $i$ finishes, once
  his rank is established he might be able to help $j$ in some way.
  He might cheer his still-running friend.  In general, we can
  consider $n$ ordered pairs $(i_r,j_r)$, $1\leq r\leq n$, with runner
  $i_1$ helping runner $j_1$, runner $i_2$ helping $j_2$, and so on.
  One simple case would be equivalence classes of
  $[n]=\{1,2,\ldots,n\}$ with each equivalence class comprising
  mutually supporting runners.  In this model, with a pair of mutually
  supporting runners only the higher-placed competitor can affect the
  other.  Each equivalence class might have a different supportive
  proclivity.

* Runner $i$ might hate runner $j$.  If $i$ finishes, he might expend
  his spare energy hindering $j$ (by throwing stones or booing , for
  example).  Note that equivalence classes of mutual hatred might
  exist.

* Also, the help or hindrance might have a finite lifetime.  Under one
  model, if $i$ finishes then he can stay helping (or hindering) $j$
  wherever $j$ places.  Alternatively, suppose $i$ places $n$th.  Then
  perhaps he can affect $j$ only place $n+1$, and after a competitor
  finishes in place $n+1$, then $i$ can exert no further influence.

* The help or hindrance discussed above might be contingent on
  placing: perhaps only the winnner (that is, first-placing
  competitor) would be able to influence other competitors.  Or maybe
  competitors with any placing could influence other competitors.

* Runners $i$ and $j$ might elect to cross the finishing line holding
  hands, with the apparent tie broken randomly.  This would be a
  symmetrical relationship.  In general, we could consider
  hand-holding equivalence classes.

* We might have a resource allocation problem, with runners having a
  certain [possibly unknown] amount of "strength" but also the ability
  to expend this strength on objectives other than ranking in the
  race.  A runner might be able to expend just enough energy to remain
  slightly ahead of a rival, reserving energy for other purposes.  If
  runner $i$ wishes to beat runner $j$, and has sufficient resources
  to do so, then runners $i$ and $j$ will finish consecutively with
  $i$ ahead of $j$.

@@

```{tikz, fig.cap = "Partial probability tree for five competitors $a-e$ with $a$ supporting $b$, hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$a\longrightarrow b$};

@@

\newcommand{\lb}{\lambda b }
\newcommand{\lc}{\lambda c }
\newcommand{\lee}{\lambda e }
\node at (6, 5)  {$\frac{\lb}{\lb+c+d+e}$};
\node at (6, 4)  {$\frac{c}{\lb+c+d+e}$};
\node at (6, 3)  {$\frac{d  }{\lb+c+d+e}$};
\node at (6, 2)  {$\frac{e  }{\lb+c+d+e}$};

\fill (7.4, 5) circle[radius=2pt];  % ab finishes; paths cde
\draw (7.4,5) -- (8.4,6);
\draw (7.4,5) -- (8.4,5);
\draw (7.4,5) -- (8.4,4);

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};
\node at (9.3, 6) {$\frac{c}{c+d+e}$};
\node at (9.3, 5) {$\frac{d  }{c+d+e}$};
\node at (9.3, 4) {$\frac{e  }{c+d+e}$};

@@

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{c}{c+e}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{e}{c+e}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{c}{c+e}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{e}{c+e}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+e}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+e}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+e}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{e}{a+b+c+e}$};

@@

\node at (11.8, 0) {$\frac{\lb}{\lb+c}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{c}{\lb+c}$};

@@

\node at (9.3, 2) {$\frac{\lb}{\lb+c+e}$};
\node at (9.3, 1) {$\frac{  c}{\lb+c+e}$};
\node at (9.3, 0) {$\frac{e  }{\lb+c+e}$};

```{tikz, fig.cap = "Partial probability tree for five competitors $a-e$ with mutually supporting  group $abc$, hyper3 approach.  If any competitor in the set {$a,b,c$} finishes, they lend their strength to the other competitors in the set who are still running", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}$};

\newcommand{\la}{\lambda a }
\newcommand{\lb}{\lambda b }
\newcommand{\lc}{\lambda c }
\newcommand{\lee}{\lambda e }
 \node at (6, 5)  {$\frac{\lb}{\lb+\lc+d+e}$};
 \node at (6, 4)  {$\frac{\lc}{\lb+\lc+d+e}$};
 \node at (6, 3)  {$\frac{d  }{\lb+\lc+d+e}$};
 \node at (6, 2)  {$\frac{e  }{\lb+\lc+d+e}$};

 \fill (7.4, 5) circle[radius=2pt];  % ab finishes; paths cde
 \draw (7.4,5) -- (8.4,6);
 \draw (7.4,5) -- (8.4,5);
 \draw (7.4,5) -- (8.4,4);

 \fill (10.2, 5) circle[radius=2pt];
 \draw (10.2, 5) -- (11.2,5.3);
 \node at (11.8, 5.3) {$\frac{\lc}{\lc+e}$};
 \node at (10.7,5.3) {$c$};

 \draw (10.2, 5) -- (11.2,4.7);
 \node at (11.8, 4.7) {$\frac{e}{\lc+e}$};
 \node at (10.7,4.7) {$e$};

 \fill (10.2, 4) circle[radius=2pt];
 \draw (10.2, 4) -- (11.2,3.9);
 \node at (11.8, 3.9) {$\frac{\lc}{\lc+e}$};
 \node at (10.7,4.1) {$c$};

 \draw (10.2, 4) -- (11.2,3.1);
 \node at (11.8, 3.1) {$\frac{e}{\lc+e}$};
 \node at (10.7,3.7) {$e$};

 \draw (4, 0) -- (5,1);
 \node at (6, 1)  {$\frac{a}{a+b+c+e}$};

 \draw (4, 0) -- (5,0);
 \node at (6, 0)  {$\frac{b}{a+b+c+e}$};

 \draw (4, 0) -- (5,-1);
 \node at (6, -1)  {$\frac{c}{a+b+c+e}$};

 \draw (4, 0) -- (5,-2);
 \node at (6, -2)  {$\frac{e}{a+b+c+e}$};

 \node at (11.8, 0) {$\frac{\lb}{\lb+\lc}$};
 \draw (10.2, -1) -- (11.2,-1);

 \node at (11.8, -1) {$\frac{c}{\lb+\lc}$};

 \draw (10.2, -2) -- (11.2,-1.6);
 \node at (11.8, -1.6) {$\frac{\la}{\la+\lc}$};

 \draw (10.2, -2) -- (11.2,-2.2);
 \node at (11.8, -2.2) {$\frac{\lc}{\la+\lc}$};

 \draw (10.2, -3) -- (11.2,-2.9);
 \node at (11.8, -2.9) {$\frac{\la}{\la+\lb}$};

 \draw (10.2, -3) -- (11.2,-3.4);
 \node at (11.8, -3.4) {$\frac{\lb}{\la+\lb}$};

@@

```{tikz, fig.cap = "Partial probability tree for five competitors $a$-$e$ with mutually supporting subsets $\\left\\lbrace a,b,c\\right\\rbrace$ [with support term $\\lambda$] and $(de)$ [with support term $\\mu$], hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (6, 5)  {$\frac{\la{b}}{\la{b}+\la{c}+d+e}$};
\node at (6, 4)  {$\frac{\la{c}}{\la{b}+\la{c}+d+e}$};
\node at (6, 3)  {$\frac{d     }{\la{b}+\la{c}+d+e}$};
\node at (6, 2)  {$\frac{e     }{\la{b}+\la{c}+d+e}$};

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};
\node at (9.3, 6) {$\frac{\la{c}}{\la{c}+d+e}$};
\node at (9.3, 5) {$\frac{d  }{\la{c}+d+e}$};
\node at (9.3, 4) {$\frac{e  }{\la{c}+d+e}$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{\ld{e}}{a+b+c+\ld{e}}$};

\node at (11.8, 0) {$\frac{\la{b}}{\la{b}+\la{c}}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{\la{c}}{\la{b}+\la{c}}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{\la{a}}{\la{a}+\la{c}}$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{\la{c}}{\la{a}+\la{c}}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{\la{a}}{\la{a}+\la{b}}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{\la{b}}{\la{a}+\la{b}}$};

\node at (9.3, 2) {$\frac{\la{b}}{\la{b}+\la{c}+e}$};
\node at (9.3, 1) {$\frac{\la{c}}{\la{b}+\la{c}+e}$};
\node at (9.3, 0) {$\frac{e     }{\la{b}+\la{c}+e}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{\la{b}}{\la{b}+\ld{e}}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{\ld{e}}{\la{b}+e}$};
